---
title: "LDA but with TidyModels"
author: "Luca Baggi"
output:
  pdf_document:
    latex_engine: xelatex
  beamer_presentation: default
  html_notebook: default
---

# Load the Libraries

```{r, message = FALSE}
library(tidymodels) # for data preprocessing
library(discrim) # for the LDA
library(MASS) # for the LDA
library(ggplot2) # for plotting the distributions
library(gridExtra)
```

# Load the data

```{r}
dir = '/Users/luca/OneDrive - Università degli Studi di Milano/uni/data-mining/lect/lect9/DFA_df.txt'

data <- read.delim(dir, sep = ',') # load the data
```
Feature selection & factorisation of the response variable:

```{r}
  df <- data %>%
  dplyr::select(-code) %>%
  mutate(y = as.factor(y))
```

I cannot standardise the data before partitioning: that's data leakage!

# Visualisations
We need to check whether the data is normally distributed.

```{r}
hist1 <- ggplot(df, aes(x1)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(x1))) +
  geom_density()

hist2 <- ggplot(df, aes(x2)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(x2))) +
  geom_density()

hist3 <- ggplot(df, aes(x3)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(x3))) +
  geom_density()
```

## Let's plot them altogether

```{r}
grid.arrange(hist1, hist2, hist3)
```

# Partition the data 

Create the split with `rsample`

```{r}
df_split <- initial_split(df, strata = y)
```

View the training and test sets:

```{r, results = FALSE}
df_split %>%
  training() %>%
  glimpse()

df_split %>%
  testing() %>%
  glimpse()
```

# Data pre-processing

## Specify the recipe

```{r recipe}
df_recipe <- training(df_split) %>%
  recipe(y ~ .) %>% # write the formula
  step_center(all_predictors()) %>% # center all predictors
  step_scale(all_predictors()) %>%
  prep()

```

## Obtain train and test data

By `juicing` and `baking` the recipe:

```{r split}
train_set <- juice(df_recipe)

test_set <- df_recipe %>%
  bake(testing(df_split))
```

We can also display train and test set:

```{r, results = FALSE}
train_set
test_set
```

## Train and test set plotting 

Much better, we can plot the train and the test set!

```{r}
norm_1 <- train_set %>%
  ggplot() +
  geom_histogram(aes(x1), alpha = 1/2, fill = 'red')
  
norm_2 <- train_set %>%
  ggplot() +
  geom_histogram(aes(x2), alpha = 1/2, fill = 'blue')

norm_3 <- train_set %>%
  ggplot() +
  geom_histogram(aes(x3), alpha = 1/2, fill = 'green')
```

And display them in a single plot:

```{r train-plots}
grid.arrange(norm_1, norm_2, norm_3)
```
X does not seem normally distributed. One should proceed with further statistical tests, such as the KS.

# LDA Model Fitting

Let's train the model:

```{r}
lda_classifier <- discrim_linear(mode = 'classification') %>%
  set_engine('MASS') %>%
  fit(y ~ ., data = train_set)
```

# Prediction

Then let's predict!

```{r}
lda_classifier %>%
  predict(test_set) %>%
  bind_cols(test_set) %>%
  glimpse()
```

`Tidymodels` will create the colums `.pred_class`: that is our prediction.

# Model Validation

```{r}
lda_classifier %>%
  predict(test_set) %>%
  bind_cols(test_set) %>%
  metrics(truth = y, estimate = .pred_class)
```
## Confusion Matrix

```{r}
lda_classifier %>%
  predict(test_set) %>%
  bind_cols(test_set) %>%
  conf_mat(y, .pred_class) %>%
  autoplot(type = 'heatmap')
```
## Let's visualise it once more

Then, let's add a layer of complications to visualise how we plotted it:


```{r}
test_set <- lda_classifier %>%
  predict(test_set) %>%
  bind_cols(test_set)

test_set %>%
  mutate(true_v_predicted = paste(test_set$y, test_set$.pred_class, sep = ',')) %>%
  ggplot() +
  geom_point(aes(x1, x2, col = true_v_predicted))
```

## Per-classifier metrics

We can obtain the metrics for each class by simply specifying a different argument in `predict`:

```{r}
lda_probs <- lda_classifier %>%
  predict(test_set, type = 'prob') %>%
  bind_cols(test_set)
```

And we can immediately plot some curves:

```{r}
lda_probs %>%
  roc_curve(y, .pred_1:.pred_3) %>%
  autoplot()
```

